---
title: "Geog533 Lab 10"
author: "Wu_Yanan"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r echo=FALSE}
if(!require(UScensus2010county)) install.county("osx")
library(UScensus2010)
library(UScensus2010county)
data(new_york.county10)
shp <- new_york.county10
plot(shp)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r echo=FALSE}
if(!require(UScensus2010county)) install.county("osx")
library(UScensus2010)
library(UScensus2010county)
library(tmap)
data(new_york.county10)
shp <- new_york.county10
qtm(shp)



```


### Question 1(c)	
How many counties in New York State?
```{r echo=FALSE}
if(!require(UScensus2010county)) install.county("osx")
library(UScensus2010)
library(UScensus2010county)
library(tmap)
data(new_york.county10)
shp <- new_york.county10
county <- shp$county
length(county)
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
if(!require(UScensus2010county)) install.county("osx")
library(UScensus2010)
library(UScensus2010county)
library(tmap)
data(new_york.county10)
shp <- new_york.county10
shp$county[shp$NAMELSAD10 == "Broome County"]

```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
library(moments)
pop <- shp$P0010001
summary(pop)
print(paste("The tota is", sum(pop)))
print(paste("The skewness is", skewness(pop)))
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(x = pop)
boxplot(x = pop)
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)

### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
if(!require(UScensus2010tract)) install.tract("osx")
data("new_york.tract10")
shp.2 <- new_york.tract10
plot(shp.2)

```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
total.2 <- shp.2$P0010001
print(paste("The total population is",sum(total.2)))
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
shp.broome<- shp.2[shp.2$county == '007',]
plot(shp.broome)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
print(paste("The total is", sum(shp.2$P0010001[shp.2$county=="007"])))
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(shp.2$P0010001)
boxplot(shp.2$P0010001)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
library(rgdal)
attribute <- shp.2[shp.2$county == "007"][1:5,]
broome.5 <- shp.2@data[shp.2@data$county == '007',][1:5,]
attribute@data
attribute@data$new.ratio <- broome.5$P0010001/sum(shp.2$P0010001)
attribute@data
writeOGR(attribute,dsn = 'output',layer = 'attribute',driver = 'ESRI Shapefile',overwrite_layer = T)
shp_write <- readOGR(dsn = 'output',layer = 'attribute')
plot(shp_write)
```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
library(raster)
library(ncdf4)
ndvi <- brick("NDVI.nc")
getwd()
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
ndvi
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
ndvi.new <- ndvi[[1:6]]
mean.ndvi <- mean(ndvi.new)
plot(mean.ndvi)
max.ndvi <- max(ndvi.new)
plot(max.ndvi)
writeRaster(mean.ndvi, filename="meanndvi.tif", overwrite=TRUE)
writeRaster(x = max.ndvi,filename = 'max_ndvi.tif',overwrite = T)
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
plot(ndvi,13:24)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r echo=FALSE}
hist(ndvi,13:28)

```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r echo=FALSE}
ndvi
plot(ndvi,7)
values <- ndvi[50,50]
plot(as.vector(value), type="b")
plot(as.vector(values), type="b")
```

